---
title: "Code_Clara"
author: "Tommaso Aicardi"
date: "2022-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo=FALSE}

#install.packages('haven')
#install.packages('dplyr')
#install.packages('leaken')
#install.packages('lsr')
#install.packages('labelled')
#install.packages('knitr')
#install.packages("kableExtra")

library(haven)
library(dplyr)
library(laeken)
library(lsr)
library(labelled)
library(knitr)
library(Hmisc)
library(kableExtra)

rper <- read_dta("rper16.dta")
carcom <- read_dta('carcom16.dta')
q16a <- read_dta('q16a.dta')

df <- merge(carcom, q16a, on='nquest')
df <- merge(df, rper, by=c('nquest','nord'))

remove('carcom','q16a','rper')

df <- data.frame(df$studio,df$y,df$area3,df$ireg)

#DF is the final dataframe containing only: Educational Attainment (STUDIO), Income per Capita(Y), Geog area(AREA3), Region(IREG).
```

# TITLE

## Inroduction

We are gonna do bla bla bla... Shiw 16, Bla bla bla...

##Exploratory Data Analysis

Here we want first to understand how our variables work, and how we can manipulate them to make them useful to our project.

```{r echo=FALSE}

kable(table(df$df.studio), caption = 'Absolute Frequencies of Educational Attainment in the Dataset', col.names = c('Educational Attainment','Frequency'))

```

We know from the [documentation](https://www.bancaditalia.it/statistiche/tematiche/indagini-famiglie-imprese/bilanci-famiglie/documentazione/documenti/2016/Quest_eng2016.pdf?language_id=1) that 6 is equal to 3y-degree. So we will put everything equal or higher than it into the same category, and everythin lower, into another cateogory.

```{r, echo=FALSE}

df$graduate <- ifelse(df$df.studio>=6,'At least Graudated','Not Graduated')



kable(table(df$graduate), caption = 'Graduated VS Not Graduated in Absolute Terms in the Dataset', col.names = c('...','Frequency'))

```
---

##Block 1 - STUDIO VS AREA3

---

##Block 2 - INCOME VS STUDIO VS IREG VS AREA3

---

##Block 3 - GEOGRAFPHICAL...
